---
title: "Rockefeller Center Energy Analysis"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

Overall energy density for the campus seperated into electric and steam sourced energy for each of the years 2019 to 2021. Steam is used for some cooling, heating and domestic hot water production. Electric covers all other services. 2019 is a re-covid year, 2020 ane 2021 show that energy use reduction were not proportional to people in the buildings. 



```{r Data Setup, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}

# Load packages 
my_packages <- c("tidyverse", "vroom" , "janitor" , "glue" , "tsibble" , "tidytext","lubridate", "fable", "tsibbledata", "ggplot2", "forecast", "tseries", "rio", "zoo", "readxl", "tsibbledata", "knitr", "kableExtra", "formattable", "scales")
invisible( lapply(my_packages, require, character.only = TRUE))

#Set up environment 
place <- "Home"  #Where are we working today. 
# place <- "work"

if (place == "Home"){setwd("C:/Users/paulr/Documents/R/RockCenterEnergyAnalysis")} else {setwd("C:/Users/prode/Documents/R/RockCenterEnergyAnalysis")}

if (!file.exists("data")) { dir.create("data")}

rm(place, my_packages ) #Clean up

options(dplyr.summarise.inform = FALSE)  # Suppress textin Knit printout. 



# Read in data

'ElecCF' <- 0.000288962  #2021 Carbon Values
'SteamCF' <- 0.00004493 #2021 Carbon Values
'IntensityLimit24' <- 0.00846 #2024 Limit
'IntensityLimit30' <- 0.00453 #2030 Limit

SteamData <- read_excel("data/Energy_Data_for_Net_Zero_Study.xlsx", sheet = "Sheet1", range = "B2:G11" , col_names = TRUE, na = "NA", col_types = NULL ) %>%  select(-5,-6)

ElectricData <- read_excel("data/Energy_Data_for_Net_Zero_Study.xlsx", sheet = "Sheet1", range = "B13:G23" , col_names = TRUE, na = "NA", col_types = NULL ) %>% select(-5,-6)

ElectricData %>% filter(Building == '1 Rock' | Building == '600 5th') %>% 
  select( -Building) %>% 
  summarise_each(funs(sum)) %>% 
  mutate(Building = "1 Rock and 600 5th") %>% 
  select(4,1,2,3) %>% rbind(ElectricData) %>% 
  filter(Building != '1 Rock' & Building != '600 5th') -> ElectricData

BuildingGSF <- read_excel("data/Energy_Data_for_Net_Zero_Study.xlsx", sheet = "Sheet1", range = "B28:C39" , col_names = TRUE, na = "NA", col_types = NULL ) %>% 
  filter(Building != '1 Rock' & Building != '600 5th')

UtilityData <- left_join(ElectricData, SteamData, by="Building") %>% left_join(BuildingGSF, by="Building")
UtilityData[is.na(UtilityData)] <- 0


# Adding NBC Electric
  All_Data <- read_excel("data/LL97 Fines Rock Center.xlsx", sheet = "NBC Energy", range = "B1:D64" , col_names = TRUE, na = "Not Available", col_types = c("guess", "guess", "numeric") ) %>% mutate(across(contains('Date'), ymd)) %>% select("Start Date", "kwh")
colnames(All_Data) <- c("Date",  "kwh")
All_Data %>% 
  group_by(year(Date)) %>% 
  summarise("Elect" = sum(kwh)) -> All_Data
colnames(All_Data) <- c("Date", "Elect")
All_Data %>% 
  filter(Date == 2019 | Date == 2020 | Date == 2021 ) -> All_Data
UtilityData[4,2] <- UtilityData[4,2] + All_Data[1,2]
UtilityData[4,3] <- UtilityData[4,3] + All_Data[2,2]
UtilityData[4,4] <- UtilityData[4,4] + All_Data[3,2]


#Campus EUI for each year 
UtilityData1 <- UtilityData 
apply(UtilityData1[2:4], 2, function(row) row * 3.4121 ) -> UtilityData1[2:4]
apply(UtilityData1[5:7], 2, function(row) row * 1194) -> UtilityData1[5:7]
UtilityData1 %>% select(`2019 kwh`, `2020 kwh`, `2021 kwh`, `2019 mlbs`, `2020 mlbs`, `2021 mlbs`, `Gross Floor Area`) %>%
summarise_each(funs(sum)) -> UtilityData1
    TotalFlArea <- UtilityData1$`Gross Floor Area`
    UtilityData1 %>% 
      select(-`Gross Floor Area`) %>% 
  gather(key = "Item", value = "Value") -> UtilityData1
  cbind(UtilityData1, str_split_fixed(UtilityData1$Item, " ", n=2)) %>% 
  select(-"Item") -> UtilityData1 
  colnames(UtilityData1)  <-  c("Value", "Year", "Unit") 
  apply(UtilityData1[3], 1, function(x) {ifelse(x == "kwh", "Elec(kBTU)", "Steam(kBTU)")}) -> UtilityData1[3]
  

  # Make an EUI plot with kBTU vales for both  electric and steam  
UtilityData2 <- UtilityData1
apply(UtilityData2[1], 1, function(row) row / TotalFlArea ) -> UtilityData2[1]
UtilityData2 %>% 
  ggplot(aes(x = Year, y = Value, fill = Unit)) +
  geom_bar(stat = "identity", position = "stack") +
  labs( title = "Energy Use Density by fuel type for Campus",
        subtitle = "Units are kBTU's per year per Gross Squar Feet") +
  xlab("Year") +
  ylab("kBTU/SF") +
  scale_fill_manual(values = c("#CC0033", "#FFFFCC"))


 ### Delta's ###
  UtilityData2 %>%
    group_by(Year) %>% 
    summarise("AnnuValues" = sum(Value)) %>% 
    mutate("BaseYear" = max(AnnuValues)) %>% 
    mutate("Delta" = AnnuValues/BaseYear) -> CampusDelta
  CampusDelta[4] <- sapply(CampusDelta[4], function(x) percent(x, accuracy=1))
```
  

A look at three years of electric and steam consumption at Rockefeller Center using just annual consumption. We capture the Covid 2019 Governmental Lock Downs imposed March of 2020 to compare with the following 2 years. All values in were converted to kBTU's from kWh of electric, and from M pounds of steam. Total energy consumed in 2020 is `r CampusDelta[2,4]` of 2019, with no little to no physical occupancy all year the relatively small reduction consumption means that consumption is a weak function of physical occupancy. Total energy consumed in 2021 is `r CampusDelta[3,4]`of 2019.

The drop in energy consumption 2019 to 2020 is due primarily the last of occupants so I posit that `r CampusDelta[2,4]` is associated with occupant activities, with the remaining to hvac, IT, lighting, vertical transportation, and miscellaneous plug loads, and building staff, offset somewhat by increased ventelation due to covid protocals. Interesting that both steam and electric dropped by similar amounts. Steam which is mostly used for heating, I would have thought would go up to make up for the occupient body heat that is lost. This leads me to believe that ventilation rates are so high and envelope losses so high that body head does not matter. 





```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}

#Make a Carbon table in the dataframe UtilityData1
UtilityData3 <- UtilityData1
  apply(UtilityData3[3], 1, function(x) {ifelse(x == "Elec(kBTU)", "Elect(tCO2e)", "Steam(tCO2e)")}) -> UtilityData3[3]
  
  
  UtilityData3 %>% 
  mutate(Carbon = ifelse(Unit == "Elec(kBTU)", Value * ElecCF / 3.4121, Value * SteamCF)) %>% ggplot(aes(x = Year, y = Carbon, fill = Unit)) +
  geom_bar(stat = "identity", position = "stack") +
  labs( title = "Carbon Emissions by fuel type",
        subtitle = "Carbon compaired to Local Law 97 limits") +
  labs(x = "Year", y = "tCO2e") +
  #guides(fill = FALSE) +
  geom_hline(aes(yintercept = TotalFlArea * IntensityLimit24), color = "blue", linetype = "dashed", size = 0.75) +
  geom_text(aes(x = 2, y = TotalFlArea * IntensityLimit24, label = "2024 Carbon Limit"), hjust = 1, vjust = 1.25 , color = "blue") +
  geom_hline(aes(yintercept = TotalFlArea * IntensityLimit30), linetype = "dashed", color = "blue", size = 0.5, alpha = 0.5) +
  geom_text(aes(x = 2.1, y = TotalFlArea * IntensityLimit30, label = "2030 Carbon Limit", alpha = 0.1), hjust = 0, vjust = -.2 , color = "blue", alpha = 0.1) +
  scale_fill_manual(values = c("#CC0033", "#FFFFCC"))
  
```

This chart show the energy consumptions in untis of carbon equilivants converted using New York City's Local Law 97 carbon factors. The dashed horzontal lings reprensent the legisated limtits. 


Energy may not be sttrickly proportion per occupient. 






```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}

#Plot Carbon Intensities versus LL97 Limits 
UtilityData1 <- UtilityData 
#apply(UtilityData1[2:4], 2, function(row) row * 3.4121 ) -> UtilityData1[2:4]
apply(UtilityData1[5:7], 2, function(row) row * 1194) -> UtilityData1[5:7]
UtilityData1[2:7]/UtilityData1$`Gross Floor Area` -> UtilityData1[2:7]
colnames(UtilityData1) <- c("Building", "2019 kwh", "2020 kwh", "2021 kwh", "2019 kBTUs", "2020 kBTUs", "2021 kBTUs", "Gross Floor Area")

UtilityData1 %>% select("Building", "2019 kwh", "2020 kwh", "2021 kwh", "2019 kBTUs", "2020 kBTUs", "2021 kBTUs") -> UtilityData1

UtilityData1 %>% 
  gather(key = "Item", value = "Value", -"Building") -> UtilityData1
  cbind(UtilityData1, str_split_fixed(UtilityData1$Item, " ", n=2)) %>% 
  select(-"Item") -> UtilityData1 
  colnames(UtilityData1)  <-  c("Building", "Value", "Year", "Unit") 
  UtilityData1 %>% 
    mutate("Carbon/SF" = ifelse(Unit == "kwh", Value * ElecCF, Value * SteamCF)) %>% 
    select(Building, Year, Unit, `Carbon/SF`) %>% 
    mutate("Carbon" = "tCO2e") -> UtilityData1
  
  UtilityData %>%   
ggplot(aes(x = Building, y = `Gross Floor Area`, fill = Building)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs( title = "Building Gross SF") +
  labs(x = "Building", y = "Gross SF") +
  theme(axis.text.x = element_text (angle = 45, vjust = 1, hjust=1)) 


UtilityData1 %>%   
ggplot(aes(x = Year, y = `Carbon/SF`, fill = Building, group = Building)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs( title = "Carbon Intensity by Building",
        subtitle = "Carbon Intensity with Local Law 97 Limtis") +
  labs(x = "Building", y = "tCO2e/sf") +
  geom_hline(aes(yintercept = IntensityLimit24), color = "blue", linetype = "dashed", size = 0.75) +
  geom_text(aes(x = 2, y = IntensityLimit24, label = "2024 Carbon Limit"), hjust = 1, vjust = 1.25 , color = "blue") +
  geom_hline(aes(yintercept = IntensityLimit30), linetype = "dashed", color = "blue", size = 0.5, alpha = 0.5) +
  geom_text(aes(x = 2, y = IntensityLimit30, label = "2030 Carbon Limit"), hjust = 0, vjust = -.2 , color = "blue", alpha = 0.01) 
  

CarbonLimits <- data.frame(Year = 2019:2021, IntensityLimit24 = 0.00846, IntensityLimit30 = 0.00453)
CarbonLimits$Year <- as.character(CarbonLimits$Year)



###### Making table for penalties ######
UtilityData1 %>% 
  left_join(BuildingGSF) %>% 
  left_join(CarbonLimits) %>% 
  mutate('tCO2e' = IntensityLimit24 - `Carbon/SF`) -> UtilityDataT

UtilityDataT %>% 
  mutate("Penality" = ifelse(tCO2e > 0, 0, (268 * tCO2e * UtilityDataT$`Gross Floor Area`))) -> UtilityDataT

UtilityDataT %>% 
  filter(Year == 2019) %>% 
  select(-Year, -Unit, -Carbon, -IntensityLimit30, -IntensityLimit24) %>% 
  group_by(Building) %>% 
  summarise( "CarbonPSF" = sum(`Carbon/SF`), "tCO2e" = sum(`tCO2e`), "Penalty" = sum(Penality))  -> UtilityDataT2

kable(UtilityDataT2, col.names = c("Building", "Carbon/GSF", "Total Carbon", "Net Penanity"), align = "lllc", caption = "Penality Tabulation", format = "simple", digits = getOption("digits"))







```

What would carbon intensities look like if all electric heating and cooling at cop =1 and then cop = 3. 









```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}

#1 Electric kWh
ElectricData %>% select(Building, `2019 kwh`, `2020 kwh`, `2021 kwh`) %>% 
  gather(key = "Item", value = "Value", -Building) %>%
  ggplot(aes(x = Building, y = Value, fill = Item, group = Item)) +
  labs( title = "Annual Electric by Building",
        subtitle = "2019 and Covid years", x = "Building", y = "kWh") +
  theme(axis.text.x = element_text (angle = 45, vjust = 1, hjust=1)) +
  geom_bar(stat = "identity", position = "dodge")
```










```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
#2 Electric kWh per GSF 
UtilityData1 <- UtilityData
apply(UtilityData1[2:7], 2, function(row) row / UtilityData1$`Gross Floor Area`) -> UtilityData1[2:7]
UtilityData1 %>% select(Building, `2019 kwh`, `2020 kwh`, `2021 kwh`) %>% 
  gather(key = "Item", value = "Value", -Building) %>% 
  ggplot(aes(x = Building, y = Value, fill = Item, group = Item)) +
  labs( title = "Annual Electric per Gross Squar Foot by Building",
        subtitle = "2019 and Covid years", x = "Building", y = "kWh/GSF") +
  theme(axis.text.x = element_text (angle = 45, vjust = 1, hjust=1)) +
  geom_bar(stat = "identity", position = "dodge")
```












```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}

#3 percent kWh reductions relative to 2019 
UtilityData1 <- UtilityData
apply(UtilityData1[2:7], 2, function(row) row / UtilityData1$`Gross Floor Area`) -> UtilityData1[2:7]
apply(UtilityData1[2:4], 2, function(row) row / UtilityData1$`2019 kwh`) -> UtilityData1[2:4]
UtilityData1 %>% select(Building, `2019 kwh`, `2020 kwh`, `2021 kwh`) %>% 
  gather(key = "Item", value = "Value", -Building) %>% 
  ggplot(aes(x = Building, y = Value, fill = Item, group = Item)) + 
  labs( title = "Percent of 2019 Electrical Consumption by Building",
        subtitle = "2019 Electric versys Covid years", x = "Building", y = "kWh/GSF % of 2019") +
  theme(axis.text.x = element_text (angle = 45, vjust = 1, hjust=1)) +
  geom_bar(stat = "identity", position = "dodge")
```
  
  
  
  
  
  
  
  
  
```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
# 4 Steam mlb 
SteamData %>% select(Building, `2019 mlbs`, `2020 mlbs`, `2021 mlbs`) %>% 
  gather(key = "Item", value = "Value", -Building) %>%
  ggplot(aes(x = Building, y = Value, fill = Item, group = Item)) +
  labs( title = "Steam Comsumption ",
        subtitle = "Steam comsumption in mLB per year and building", x = "Building", y = "Mlb") +
  theme(axis.text.x = element_text (angle = 45, vjust = 1, hjust=1)) +
  geom_bar(stat = "identity", position = "dodge")

```








```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}

#5 steam mlb per gsf 
UtilityData1 <- UtilityData
apply(UtilityData1[2:7], 2, function(row) row / UtilityData1$`Gross Floor Area`) -> UtilityData1[2:7]
UtilityData1 %>% select(Building, `2019 mlbs`, `2020 mlbs`, `2021 mlbs`) %>% 
  gather(key = "Item", value = "Value", -Building) %>% 
  ggplot(aes(x = Building, y = Value, fill = Item, group = Item)) +
  labs( title = "Steam Comsumption per GSF ",
        subtitle = "Steam comsumption in mLB per GSF year and building", x = "Building", y = "Mlb/GSF") +
  theme(axis.text.x = element_text (angle = 45, vjust = 1, hjust=1)) +
  geom_bar(stat = "identity", position = "dodge")
  
```









```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}

#6 percent down from 2019
UtilityData1 <- UtilityData
apply(UtilityData1[2:7], 2, function(row) row / UtilityData1$`Gross Floor Area`) -> UtilityData1[2:7]
apply(UtilityData1[5:7], 2, function(row) row / UtilityData1$`2019 mlbs`) -> UtilityData1[5:7]
UtilityData1 %>% select(Building, `2019 mlbs`, `2020 mlbs`, `2021 mlbs`) %>% 
  gather(key = "Item", value = "Value", -Building) %>% 
  ggplot(aes(x = Building, y = Value, fill = Item, group = Item)) +
  labs( title = "Percent of 2019 Steam Consumption by Building ",
        subtitle = "2019 Steam versys Covid years", x = "Building", y = "Mlb/GSF % of 2019") +
  theme(axis.text.x = element_text (angle = 45, vjust = 1, hjust=1)) +
  geom_bar(stat = "identity", position = "dodge")

```









```{r  echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}


EPA_data <- read_excel("data/EPA_Annual_Energy_Use_By_Meter.xlsx", range = "A6:BB155" , col_names = FALSE, na = "NA", col_types = NULL )

EPA_data <- read_excel("data/EPA_Annual_Energy_Use_By_Meter.xlsx", skip = 4)




Turnstile_data <- read.csv("data/RC_Occunpancy_Data_Details_data.csv") %>% 
  select(Property, Date, Total.Entrants)

Turnstile_data$Date <- mdy(Turnstile_data$Date)

Turnstile_data %>%   
group_by(Property, "Year" = year(Date)) %>%  #"Month" = month(Date)) %>% 
  summarise(Population = sum(Total.Entrants)) -> Turnstile_data1
Turnstile_data1$Year <- as.factor(Turnstile_data1$Year)

Turnstile_data1 %>% 
  ggplot(aes(x = Property, y = Population, group = Year, fill = Year)) +
   labs( title = "Turnstyle Counts by Month",
        subtitle = "Morning Counts", x = "Building", y = "Entrants") +
  theme(axis.text.x = element_text (angle = 45, vjust = 1, hjust=1)) +
  geom_bar(stat = "identity", position = "dodge")
  




Turnstile_data %>%   
group_by("Year" = year(Date)) %>%   
  summarise(Population = sum(Total.Entrants)) -> Turnstile_data2
Turnstile_data2$Year <- as.factor(Turnstile_data2$Year)


  

```
